Variant calling with transformers

### 2022-06-28 manuscript/manuscript.tex
Added short variant context section and plot

>>>>

Model input, derived from aligned reads, includes information about nearby non-reference bases and other context information such as base qualities. We were interested to see how model variant calling was informed by or correlated to aligned read information near a prospective variant. Prospective SNVs with a VAF much lower than the roughly 50\% expected for a nominal heterozygous variant are less likely to be true variants and are less likely to be called than those with higher VAF. This model decision process may involve the use of local context information. For example, the model may detect many low VAF prospective variants in a local region. In many cases this can be related to a difficult to sequence region that presents numerous false positive prospective variants. Through training, the model may learn to reduce variant call probability in such regions. Base quality scores, a local context parameter that can also indicate incorrect non-reference base calls, may also play an important roll in informing the model that a prospective variant call is incorrect.

To explore the transformer models use of these local context data for prospective low VAF variants, we evaluated low VAF variant call probability relative to the total number of nearby low VAF prospective variants. Low versus High base quality was included to determine if it was also correlated to call probability and possibly used in the model decision process. Figure XXX shows three different low VAF ranges for prospective variants and what fraction of these variants were called by the model compared to the number of other nearby low VAF variants and also compared to base quality score. The results show a strong trend between low local prospective variant population and higher probability of being called by the model. Base quality scores, on the other hand, were not strongly linked to call probability or to the local population of prospective variants. This indicates that to some extent the model uses local non-reference base information to determine variant calls.
